---
title: "Compute correlation analysis in the different clusters (1000 random)"
output:
    html_document:
        toc: TRUE
meta:
    exp_id: 20211104_01_lncRNAsExpressionCorrelation_1000
    type: analyse
    modules:
        - R/4.1
---

# Environment setup

```{r setup}

library(tidyverse)
library(ggpubr)
library(rstatix)

theme_set(theme_bw())

```

## Max lncRNA/coding correlation all genes

```{r}

coexp_summary <-
    read_tsv("data/clusters/random_1000/coexp_lncRNA_max_corr.tsv")

coexp_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 3)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()


ttest_summary <-
    t_test(corr ~ cluster_type, data=ungroup(coexp_summary)) %>%
    add_xy_position(x="cluster_type", step.increase = .7)
graph <-
    coexp_summary %>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/coexp_violin_max_lncRNA_correlation.pdf",
       plot = graph, width = 4, height = 5)
```

## Max lncRNA/coding correlation no poorly detected genes

```{r}

coexp_summary <-
    read_tsv("data/clusters/random_1000/coexp_lncRNA_max_corr_no_poorly_detected.tsv")

coexp_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 3)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()


ttest_summary <-
    t_test(corr ~ cluster_type, data=ungroup(coexp_summary)) %>%
    add_xy_position(x="cluster_type", step.increase = .5)

coexp_summary$cluster_type <-
  factor(coexp_summary$cluster_type,
         c("random","full_coexp","coexp"))

graph <-
    coexp_summary%>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/clusters/1000/coexp_violin_max_lncRNA_correlation_no_poorly_detected.pdf",
       plot = graph, width = 5, height = 4)
```

##  Median cluster coding gene coexpression

```{r}

coexp_summary <-
    read_tsv("data/clusters/random_1000/coexp_coding_corr.tsv") %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

coexp_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(median_corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 6)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

ttest_summary <-
    coexp_summary %>%
    group_by(cluster_type) %>%
    t_test(median_corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue() %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

graph <-
    coexp_summary %>%
    ggplot(aes(x=group, y=median_corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")
print(graph)

ggsave("out/coexp_violin_median_cluster_correlation.pdf",
       plot = graph, width = 6, height = 5)
```

## Median coding gene co-expression no poorly detected genes

```{r}

coexp_summary <-
    read_tsv("data/clusters/random_1000/coexp_coding_corr_no_poorly_detected.tsv") %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

coexp_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(median_corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 6)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

ttest_summary <-
    coexp_summary %>%
    group_by(cluster_type) %>%
    t_test(median_corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue(method = "fdr") %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

coexp_summary$group <-
  factor(coexp_summary$group,
         c("randomFALSE","randomTRUE","full_coexpFALSE","full_coexpTRUE","coexpFALSE","coexpTRUE"))


graph <-
    coexp_summary %>%
    ggplot(aes(x=group, y=median_corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")+
  theme(axis.text.x  = element_text(angle=45, vjust=1, hjust=1, size=10))
print(graph)

ggsave("out/clusters/1000/coexp_violin_median_cluster_correlation_no_poorly_detected.pdf",
       plot = graph, width = 6, height = 5)

```

# plantiSMASH clusters

```{r plantiSMASHcluster_analysis}

MRN_cluster <- "Cluster 39_0"

```

## Max lncRNA/coding gene correlation

```{r}

plantiSMASH_summary <-
    read_tsv("data/clusters/random_1000/plantiSMASH_lncRNA_max_corr.tsv")

plantiSMASH_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 2)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

MRN_cluster_info <-
    plantiSMASH_summary %>%
    filter(cluster_name == MRN_cluster) %>%
    mutate(cluster_name="marneral cluster")

graph <-
    plantiSMASH_summary %>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    geom_point(data=MRN_cluster_info, color="red") +
    geom_text(data=MRN_cluster_info, color="red", aes(label=cluster_name),
              vjust=-1) +
    stat_compare_means(method="t.test",
                       comparisons=list(c("random", "plantiSMASH")),
                       label="p.value",
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/plantiSMASH_violin_max_lncRNA_correlation.pdf",
       plot = graph, width = 3, height = 2.5)
```

## Max lncRNA/coding gene correlation non poorly detected genes

```{r}

plantiSMASH_summary <-
    read_tsv("data/clusters/random_1000/plantiSMASH_lncRNA_max_corr_no_poorly_detected.tsv")

plantiSMASH_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 2)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

MRN_cluster_info <-
    plantiSMASH_summary %>%
    filter(cluster_name == MRN_cluster) %>%
    mutate(cluster_name="marneral cluster")


plantiSMASH_summary$cluster_type <-
  factor(plantiSMASH_summary$cluster_type,
         c("random","plantiSMASH"))

graph <-
    plantiSMASH_summary %>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    geom_point(data=MRN_cluster_info, color="red") +
    geom_text(data=MRN_cluster_info, color="red", aes(label=cluster_name),
              vjust=-1) +
    stat_compare_means(method = "t.test",
                       comparisons=list(c("random", "plantiSMASH")),
                       label="p.value",
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/clusters/1000/plantiSMASH_violin_max_lncRNA_correlation_no_poorly_detected.pdf",
       plot = graph, width = 3, height = 4)
```

## Median cluster coding gene correlation

```{r}

plantiSMASH_summary <-
    read_tsv("data/clusters/random_1000/plantiSMASH_coding_corr.tsv") %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

plantiSMASH_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(median_corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 4)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

MRN_cluster_info <-
    plantiSMASH_summary %>%
    filter(cluster_name == MRN_cluster) %>%
    mutate(cluster_name="marneral\ncluster")

ttest_summary <-
    plantiSMASH_summary %>%
    group_by(cluster_type) %>%
    t_test(median_corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue() %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

graph <-
    plantiSMASH_summary %>%
    ggplot(aes(x=group, y=median_corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    geom_point(data=MRN_cluster_info, color="red") +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")
print(graph)

ggsave("out/plantiSMASH_violin_median_cluster_correlation.pdf",
       plot = graph, width = 4, height = 5)
```

## Median cluster coding gene correlation no poorly detected genes

```{r}

plantiSMASH_summary <-
    read_tsv("data/clusters/random_1000/plantiSMASH_coding_corr_no_poorly_detected.tsv") %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

plantiSMASH_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(median_corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 4)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

MRN_cluster_info <-
    plantiSMASH_summary %>%
    filter(cluster_name == MRN_cluster) %>%
    mutate(cluster_name="marneral\ncluster")

ttest_summary <-
    plantiSMASH_summary %>%
    group_by(cluster_type) %>%
    t_test(median_corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue(method = "fdr") %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)


plantiSMASH_summary$group <-
  factor(plantiSMASH_summary$group,
         c("randomFALSE","randomTRUE","plantiSMASHFALSE","plantiSMASHTRUE"))

graph <-
    plantiSMASH_summary %>%
    ggplot(aes(x=group, y=median_corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
        geom_point(data=MRN_cluster_info, color="red") +
    scale_fill_brewer(palette= "Set1") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")+
    ylim(-0.5,1)
print(graph)

ggsave("out/clusters/1000/plantiSMASH_violin_median_cluster_correlation_no_poorly_detected.pdf",
       plot = graph, width = 5, height = 4)
```

