---
title: "Compute correlation analysis in the different clusters"
output:
    html_document:
        toc: TRUE
meta:
    exp_id: 20211029_07_lncRNAsExpressionCorrelation
    type: analyse
    modules:
        - R/4.1
---

# Environment setup

```{r setup}

library(tidyverse)
library(ggpubr)
library(rstatix)

theme_set(theme_bw())

```

# Coexp clusters

```{r coexp_cluster_analysis}

coexp_corr_info <-
    read_tsv("data/clusters/coexp_corr_info.tsv.gz")

coexp_expression <-
    read_tsv("data/clusters/coexp_expression_summary.tsv")

coexp_poorly_expressed <-
    coexp_expression %>%
    filter(detected_libraries < 20) %>%
    pull(gene_id)

```

## Max lncRNA/coding correlation all genes

```{r}

coexp_summary <-
    coexp_corr_info %>%
    filter(lncRNA_cluster, gene_type == "coding/lncRNA") %>%
    group_by(cluster_name, cluster_type) %>%
    summarise(corr = max(corr))

coexp_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 3)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()


ttest_summary <-
    t_test(corr ~ cluster_type, data=ungroup(coexp_summary)) %>%
    add_xy_position(x="cluster_type", step.increase = .7)

coexp_summary$cluster_type <- factor(coexp_summary$cluster_type, levels=c("random","full_coexp","coexp"))

graph <-
    coexp_summary %>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/clusters/final/coexp_violin_max_lncRNA_correlation.pdf",
       plot = graph, width = 4, height = 5)
```

## Max lncRNA/coding correlation no poorly detected genes

```{r}

coexp_summary <-
    coexp_corr_info %>%
    filter(lncRNA_cluster, gene_type == "coding/lncRNA",
           ! gene_id_1 %in% coexp_poorly_expressed,
           ! gene_id_2 %in% coexp_poorly_expressed) %>%
    group_by(cluster_name, cluster_type) %>%
    summarise(corr = max(corr))

coexp_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 3)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()


ttest_summary <-
    t_test(corr ~ cluster_type, data=ungroup(coexp_summary)) %>%
    add_xy_position(x="cluster_type", step.increase = .5)

coexp_summary$cluster_type <- factor(coexp_summary$cluster_type, levels=c("random","full_coexp","coexp"))

graph <-
    coexp_summary%>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/clusters/final/coexp_violin_max_lncRNA_correlation_no_poorly_detected.pdf",
       plot = graph, width = 4, height = 5)
```

##  Median cluster coding gene coexpression

```{r}

coexp_summary <-
    coexp_corr_info %>%
    filter(gene_type == "coding") %>%
    group_by(cluster_name, cluster_type, lncRNA_cluster) %>%
    summarise(corr = median(corr)) %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

coexp_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 6)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

ttest_summary <-
    coexp_summary %>%
    group_by(cluster_type) %>%
    t_test(corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue() %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

graph <-
    coexp_summary %>%
    ggplot(aes(x=group, y=corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")
print(graph)

ggsave("out/clusters/final/coexp_violin_median_cluster_correlation.pdf",
       plot = graph, width = 6, height = 5)
```

## Median coding gene co-expression no poorly detected genes

```{r}

coexp_summary <-
    coexp_corr_info %>%
    filter(gene_type == "coding",
           ! gene_id_1 %in% coexp_poorly_expressed,
           ! gene_id_2 %in% coexp_poorly_expressed) %>%
    group_by(cluster_name, cluster_type, lncRNA_cluster) %>%
    summarise(corr = median(corr)) %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

coexp_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 6)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

ttest_summary <-
    coexp_summary %>%
    group_by(cluster_type) %>%
    t_test(corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue() %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

coexp_summary$group <- factor(coexp_summary$group, levels=c("randomFALSE","randomTRUE", "full_coexpFALSE","full_coexpTRUE","coexpFALSE","coexpTRUE"))


graph <-
    coexp_summary %>%
    ggplot(aes(x=group, y=corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")+
    theme (axis.text.x  = element_text(angle=90, vjust=.5, size=10))
print(graph)

ggsave("out/clusters/final/coexp_violin_median_cluster_correlation_no_poorly_detected.pdf",
       plot = graph, width = 6, height = 5)

```

# Low coexp clusters

```{r low_coexp_cluster_analysis}

low_coexp_corr_info <-
    read_tsv("in/20211030_02_lncRNAsExpressionCorrelationLowCoexpRandom/low_coexp_corr_info.tsv.gz")

low_coexp_expression <-
    read_tsv("in/20211030_02_lncRNAsExpressionCorrelationLowCoexpRandom/low_coexp_expression_summary.tsv")

low_coexp_poorly_expressed <-
    low_coexp_expression %>%
    filter(detected_libraries < 20) %>%
    pull(gene_id)

```

## Max lncRNA/coding correlation all genes

```{r}

low_coexp_summary <-
    low_coexp_corr_info %>%
    filter(lncRNA_cluster, gene_type == "coding/lncRNA") %>%
    group_by(cluster_name, cluster_type) %>%
    summarise(corr = max(corr))

low_coexp_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 3)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()


ttest_summary <-
    t_test(corr ~ cluster_type, data=ungroup(low_coexp_summary)) %>%
    add_xy_position(x="cluster_type", step.increase = .8)

graph <-
    low_coexp_summary %>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/low_coexp_violin_max_lncRNA_correlation.pdf",
       plot = graph, width = 4, height = 5)
```

## Max lncRNA/coding correlation no poorly detected genes

```{r}

low_coexp_summary <-
    low_coexp_corr_info %>%
    filter(lncRNA_cluster, gene_type == "coding/lncRNA",
           ! gene_id_1 %in% low_coexp_poorly_expressed,
           ! gene_id_2 %in% low_coexp_poorly_expressed) %>%
    group_by(cluster_name, cluster_type) %>%
    summarise(corr = max(corr))

low_coexp_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 3)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

ttest_summary <-
    t_test(corr ~ cluster_type, data=ungroup(low_coexp_summary)) %>%
    add_xy_position(x="cluster_type", step.increase = .8)

graph <-
    low_coexp_summary%>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/low_coexp_violin_max_lncRNA_correlation_no_poorly_detected.pdf",
       plot = graph, width = 4, height = 5)
```

##  Median cluster coding gene coexpression

```{r}

low_coexp_summary <-
    low_coexp_corr_info %>%
    filter(gene_type == "coding") %>%
    group_by(cluster_name, cluster_type, lncRNA_cluster) %>%
    summarise(corr = median(corr)) %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

low_coexp_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 6)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

ttest_summary <-
    low_coexp_summary %>%
    group_by(cluster_type) %>%
    t_test(corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue() %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

graph <-
    low_coexp_summary %>%
    ggplot(aes(x=group, y=corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")
print(graph)

ggsave("out/low_coexp_violin_median_cluster_correlation.pdf",
       plot = graph, width = 6, height = 5)
```

## Median coding gene co-expression no poorly detected genes

```{r}

low_coexp_summary <-
    low_coexp_corr_info %>%
    filter(gene_type == "coding",
           ! gene_id_1 %in% low_coexp_poorly_expressed,
           ! gene_id_2 %in% low_coexp_poorly_expressed) %>%
    group_by(cluster_name, cluster_type, lncRNA_cluster) %>%
    summarise(corr = median(corr)) %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

low_coexp_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 6)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

ttest_summary <-
    low_coexp_summary %>%
    group_by(cluster_type) %>%
    t_test(corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue() %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

graph <-
    low_coexp_summary %>%
    ggplot(aes(x=group, y=corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")
print(graph)

ggsave("out/low_coexp_violin_median_cluster_correlation_no_poorly_detected.pdf",
       plot = graph, width = 6, height = 5)

```

# plantiSMASH clusters

```{r plantiSMASHcluster_analysis}

plantiSMASH_corr_info <-
    read_tsv("data/clusters/plantiSMASH_corr_info.tsv.gz")

plantiSMASH_expression <-
    read_tsv("data/clusters/plantiSMASH_expression_summary.tsv")

plantiSMASH_poorly_expressed <-
    plantiSMASH_expression %>%
    filter(detected_libraries < 20) %>%
    pull(gene_id)

MRN_cluster <-
    plantiSMASH_corr_info %>%
    filter(gene_id_1 == "AT5G00580" | gene_id_2 == "AT5G00580") %>%
    pull(cluster_name) %>%
    unique


```

## Max lncRNA/coding gene correlation

```{r}

plantiSMASH_summary <-
    plantiSMASH_corr_info %>%
    filter(lncRNA_cluster, gene_type == "coding/lncRNA") %>%
    group_by(cluster_name, cluster_type) %>%
    summarise(corr = max(corr))

plantiSMASH_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 2)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

MRN_cluster_info <-
    plantiSMASH_summary %>%
    filter(cluster_name == MRN_cluster) %>%
    mutate(cluster_name="marneral cluster")

graph <-
    plantiSMASH_summary %>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    geom_point(data=MRN_cluster_info, color="red") +
    geom_text(data=MRN_cluster_info, color="red", aes(label=cluster_name),
              vjust=-1) +
    stat_compare_means(method="t.test",
                       comparisons=list(c("random", "plantiSMASH")),
                       label="p.value",
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/clusters/final/plantiSMASH_violin_max_lncRNA_correlation.pdf",
       plot = graph, width = 3, height = 5)
```

## Max lncRNA/coding gene correlation non poorly detected genes

```{r}

plantiSMASH_summary <-
    plantiSMASH_corr_info %>%
    filter(lncRNA_cluster, gene_type == "coding/lncRNA",
           ! gene_id_1 %in% plantiSMASH_poorly_expressed,
           ! gene_id_2 %in% plantiSMASH_poorly_expressed) %>%
    group_by(cluster_name, cluster_type) %>%
    summarise(corr = max(corr))

plantiSMASH_summary %>%
    group_by(cluster_type) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 2)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()


MRN_cluster_info <-
    plantiSMASH_summary %>%
    filter(cluster_name == MRN_cluster) %>%
    mutate(cluster_name="marneral cluster")

graph <-
    plantiSMASH_summary %>%
    ggplot(aes(x=cluster_type, y=corr)) +
    geom_violin() +
    geom_boxplot(width = .2, outlier.shape = NA) +
    geom_point(data=MRN_cluster_info, color="red") +
    geom_text(data=MRN_cluster_info, color="red", aes(label=cluster_name),
              vjust=-1) +
    stat_compare_means(method="t.test",
                       comparisons=list(c("random", "plantiSMASH")),
                       label="p.value",
                       bracket.size=.5) +
    xlab("") + 
    ylab("maximum lncRNA correlation")
print(graph)

ggsave("out/clusters/final/plantiSMASH_violin_max_lncRNA_correlation_no_poorly_detected.pdf",
       plot = graph, width = 3, height = 5)
```

## Median cluster coding gene correlation

```{r}

plantiSMASH_summary <-
    plantiSMASH_corr_info %>%
    filter(gene_type == "coding") %>%
    group_by(cluster_name, cluster_type, lncRNA_cluster) %>%
    summarise(corr = median(corr)) %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

plantiSMASH_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 4)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

MRN_cluster_info <-
    plantiSMASH_summary %>%
    filter(cluster_name == MRN_cluster) %>%
    mutate(cluster_name="marneral\ncluster")

ttest_summary <-
    plantiSMASH_summary %>%
    group_by(cluster_type) %>%
    t_test(corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue() %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

graph <-
    plantiSMASH_summary %>%
    ggplot(aes(x=group, y=corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")
print(graph)

ggsave("out/clusters/final/plantiSMASH_violin_median_cluster_correlation.pdf",
       plot = graph, width = 4, height = 5)
```

## Median cluster coding gene correlation no poorly detected genes

```{r}

plantiSMASH_summary <-
    plantiSMASH_corr_info %>%
    filter(gene_type == "coding",
           ! gene_id_1 %in% plantiSMASH_poorly_expressed,
           ! gene_id_2 %in% plantiSMASH_poorly_expressed) %>%
    group_by(cluster_name, cluster_type, lncRNA_cluster) %>%
    summarise(corr = median(corr)) %>%
    mutate(group=str_c(cluster_type, lncRNA_cluster))

plantiSMASH_summary %>%
    group_by(cluster_type, lncRNA_cluster) %>%
    summarise(value=quantile(corr)) %>%
    bind_cols(quantile=rep(c(0,.25,.5,.75,1), 4)) %>%
    pivot_wider(names_from="quantile", values_from="value") %>%
    knitr::kable()

MRN_cluster_info <-
    plantiSMASH_summary %>%
    filter(cluster_name == MRN_cluster) %>%
    mutate(cluster_name="marneral\ncluster")

ttest_summary <-
    plantiSMASH_summary %>%
    group_by(cluster_type) %>%
    t_test(corr ~ group) %>%
    # since test on group we need to apply multitest correction afterwards
    adjust_pvalue() %>%
    add_significance() %>%
    add_xy_position(x="group", step.increase = .7)

graph <-
    plantiSMASH_summary %>%
    ggplot(aes(x=group, y=corr)) +
    geom_violin(position=position_dodge(.9), scale="width", aes(fill=lncRNA_cluster)) +
    geom_boxplot(width = .2, position=position_dodge(.9), outlier.shape = NA, aes(fill=lncRNA_cluster)) +
    scale_fill_brewer(palette= "Accent") +
    stat_pvalue_manual(ttest_summary, label = "p.adj", 
                       bracket.size=.5) +
    xlab("") + 
    ylab("median lncRNA correlation")
print(graph)

ggsave("out/clusters/final/plantiSMASH_violin_median_cluster_correlation_no_poorly_detected.pdf",
       plot = graph, width = 4, height = 5)
```

