---
title: "Seed germination analysis (ABA treated MHAL)"
author: "Thomas Blein"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: cite.bib
output:
    html_document:
        toc: true
        toc_float: true
        number_sections: true
link-citations: yes
---

A review to how to analyse censored data in biology was done by
@doi:10.1111/aab.12477 . Seed germination are censored data, ie we monitored the
seeds that germinate before a certain date. So we have time intervals. Classical
statistics (ie ANOVA) are not working nicely to analyse this kind of data. None
linear regression models need to be applied as for survival analysis, except
that the curve is inverted.

Several model can be fitted to germination data. The package `drc` propose two
of them :

- the classical [logistic
  distribution](https://en.wikipedia.org/wiki/Logistic_distribution), which is
  symmetric at the inflection point (T50, or time of 50% of germination).
- [Weibull distributions](https://en.wikipedia.org/wiki/Weibull_distribution),
  which is not symmetric at T50. Two type of parametrisation exist for the
  Weibull function that are referred in `drc` package as `weibull1` and
  `weibull2`.

The method used here was adapted from method proposed on [Statistical Analysis of Agricultural
Experiments using R](https://rstats4ag.org/), [chapter 14 Seed
Germination](https://rstats4ag.org/seed-germination.html) and [Analysing seed
germination data with R:
a tutorial](https://www.statforbiology.com/seedgermination/index.html_document)

Load the packages and set the default theme.

```{r setup}

library("tidyverse")
library("multcompView")
library("drc")

theme_set(theme_bw())

```

# Data preparation

Load the raw data from TSV file.

```{r load_data}

raw_data <-
    read_tsv("./count_ABA_12.tsv", col_type=cols()) %>%
    mutate(CumGerm=total_nb * radicule_out / 100,
           concentration=factor(concentration),
           plate_nb=factor(plate_nb),
           genotype=genotype %>% tolower %>% factor)

genotype_info <-
    tibble(genotype_name=c("Col", "mrn1", "pro T-DNA", "RNAi 3.8", "RNAi 9.2",
                           "35S:MRN1")) %>%
    mutate(genotype=str_replace_all(genotype_name, "[:. -]", "") %>%
                tolower %>%
                str_replace("tdna", ""),
            genotype_name=factor(genotype_name, levels=genotype_name))

```

Format the data to be analysed with the `drc` package.

- Filter the data
- If cumulative count > total count set as total count If the cumulative is
  below the previous cumulative set it as the previous one (CumGerm at 2PM
  < CumGerm at 2AM set it to CumGerm at 2AM).
- Transform counts them as by day progress and not cumulative Transform the time
  name to the number of hour, discrete to continuous scale.
- The process require to have the upper and lower bound set for the time: how
  many seeds germinate since last counting time. So for time 0 the upper bound
  is set at 1h and lower bound of 0h. For 1AM, lower bound to 1h and upper bound
  to 1AM (24h). For 1PM, lower bound set to 1AM (24h) and upper bound set to 1PM
  (32h). Etc...
- Add and extra time to take into the seeds that did not
  germinated : lower bound set to 3PM (80h) and upper bound to Inf.
- Remove the `:` and `.` from the genotype names to prevent any strange
  behavior (`:` is used as a separating field in the analysis).

```{r data_preparation}

# time between plant out and midnight
time_shift <- 24 - 10.5 # out at 10h30

# the measurements by themself
formated_measurements <-
    raw_data %>%
    dplyr::select(concentration, genotype, plate_nb, day, total_nb, CumGerm) %>%
    # filter the time
    filter(!is.na(plate_nb),
           !is.na(CumGerm),
           day %in% c("0", "1AM", "1PM", "2AM", "2PM", "3AM", "3PM")) %>%
    # Arrange the data so the different times are on following lines per plate
    # to be able to calculate daily germination
    arrange(concentration, genotype, day, plate_nb) %>%
    group_by(concentration, genotype, plate_nb) %>%
    mutate(
           # Correct cumulative germination
           # more germinated seeds than total number of seeds ;)
           CumGerm=if_else(CumGerm > total_nb, total_nb, CumGerm),
           # more seeds on previous day.
           CumGerm=if_else(CumGerm < lag(CumGerm), lag(CumGerm), CumGerm),
           CumGerm=if_else(is.na(CumGerm), 0, CumGerm),
           # Calculate the daily germination
           DailyGerm=CumGerm - lag(CumGerm),
           # Set the daily germination of time point 0 to 0
           DailyGerm=if_else(is.na(DailyGerm), 0, DailyGerm)) %>%
    ungroup %>%
    mutate(
           # Clean genotype name
           genotype=str_replace_all(genotype, "[:.]", ""),
           # Set the lower bounds of the time intervals
           timeBef=case_when(
                   day == "0" ~ 0,
                   day == "1AM" ~ 1,
                   day == "1PM" ~ 10.5 + time_shift,
                   day == "2AM" ~ 18.5 + time_shift,
                   day == "2PM" ~ 10.5 + 24 + time_shift,
                   day == "3AM" ~ 18.5 + 24 + time_shift,
                   day == "3PM" ~ 10.5 + 48 + time_shift
                   ),
           # Set the upper bounds of the time intervals
           timeAf=case_when(
                   day == "0" ~ 1,
                   day == "1AM" ~ 10.5 + time_shift,
                   day == "1PM" ~ 18.5 + time_shift,
                   day == "2AM" ~ 10.5 + 24 + time_shift,
                   day == "2PM" ~ 18.5 + 24 + time_shift,
                   day == "3AM" ~ 10.5 + 48 + time_shift,
                   day == "3PM" ~ 18.5 + 48 + time_shift
                   ),
           # compute a unique plate identification
           plate_id=str_c(concentration, genotype, plate_nb) %>% as.factor %>% as.numeric)

# The plant that do not germinate
ungerminated_data <-
    formated_measurements %>%
    # Keep only last time point
    filter(day == "3PM") %>%
    # remove unneeded columns that will be generated
    dplyr::select(-(DailyGerm:timeAf)) %>%
    unique %>%
    mutate(
           # The daily germination seeds (the seeds that did not germinated) is
           # the total number of seeds minus the seeds that already germinated
           DailyGerm=total_nb - CumGerm,
           # The cumulative number of seed is all the seeds
           CumGerm=total_nb,
           # Set the bound of the time interval
           timeBef = 18.5 + 48 + time_shift,
           timeAf = Inf
           )

# fusion of the data
formated_data <-
    bind_rows(formated_measurements, ungerminated_data) %>%
    arrange(plate_id, timeBef) %>%
    group_by(plate_id) %>%
    # Generate a line id that should look like:
    # 1    for first plate, zero time point
    # 1.1  for first plate, time point 1
    # 1.2  ...
    # ...
    # 2    for second plate, zero time point
    # 2.1  for second plate, time point 1
    # 1.2  ...
    # ...
    mutate(rownames=str_c(plate_id, ".", row_number() - 1) %>%
                str_replace("\\.0", "")
           ) %>%
    # Remove unneeded columns
    dplyr::select(-day, -plate_nb) %>%
    # transform it as data.frame since `dcr` package do not work with tibble
    as.data.frame %>%
    # set the row names with the constructed ID
    column_to_rownames("rownames")

formated_data %>%
    head(10)

```

# 0.5 ABA

Parameters significance in the models:

- `Max`: the upper limit being the maximal germination, with an upper limit set
  at 1 (or 100% of the seed);
- `T50` is the time taken to reach 50% germination relative to the upper limit.
- `Slope` could give an indication of how quickly seed germination reaches 
  completion once the process has begun.
- `T5` or `T95` (the time required for 5 and 95 percent of the seeds to
  germinate, respectively) can also be estimated in relation to the upper limit.

## Model fitting

Compare and model with 3 parameters and 2 parameters (maximum set to 1).

```{r model_comparison}

mod_05_LL3 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             curveid = genotype,
             data = formated_data,
             fct = LL.3(names=c("Slope", "Max", "T50")),
             type = "event", subset=c(concentration == 0.5))
plot(mod_05_LL3, log = "", legendPos = c(80, .5),
     ylab="Proportion of germinated seed", xlab="Time (h)", main="0.5 ABA")
summary(mod_05_LL3)

mod_05_LL2 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             curveid = genotype,
             data = formated_data,
             fct = LL.2(names=c("Slope", "T50")),
             type = "event", subset=c(concentration == 0.5))
# Add this model to the plot in red
plot(mod_05_LL2, log = "", type = "none", legendPos = c(80, .5), col=c(2), add=TRUE)
summary(mod_05_LL2)

anova(mod_05_LL3, mod_05_LL2, test = "Chisq")

```

There is difference between the two model. The main difference being the
maximum, we can speculate that the max is not 1 (not all the seed germinate)

## Global genotype effect

We can then wonder if there is a global effect of the genotype. We compare
a model where there a genotype effect with one were there is none.

```{r modNull_05_LL3}

modNull_05_LL3 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             data = formated_data,
             fct = LL.3(names=c("Slope", "Max", "T50")),
             type = "event", subset=c(concentration == 0.5),
             upperl=c(NA,1,NA))
plot(mod_05_LL3, log = "", legendPos = c(80, .5),
     ylab="Proportion of germinated seed", xlab="Time (h)", main="0.5 ABA")
plot(modNull_05_LL3, log = "", type = "none", legendPos = c(80, .5), col=c(3), add=TRUE)

anova(mod_05_LL3, modNull_05_LL3, test = "Chisq")
```
In that case we specify the maximum value for `Max` to 1 otherwise the model
fitting crash. We do not put it on the previous LL3 model for same raison.

```{r modNull_05_LL2}

modNull_05_LL2 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             data = formated_data,
             fct = LL.2(names=c("Slope", "T50")),
             type = "event", subset=c(concentration == 0.5))
plot(mod_05_LL2, log = "", legendPos = c(80, .5),
     ylab="Proportion of germinated seed", xlab="Time (h)", main="0.5 ABA")
plot(modNull_05_LL2, log = "", type = "none", legendPos = c(80, .5), col=c(4), add=TRUE)

anova(mod_05_LL2, modNull_05_LL2, test = "Chisq")

```

Strong differences for both null models.

## Model comparison

Which one the best? Use the AIC and BIC criteria:

```{r model_choice}

AIC(mod_05_LL3, mod_05_LL2, modNull_05_LL3, modNull_05_LL2)
BIC(mod_05_LL3, mod_05_LL2, modNull_05_LL3, modNull_05_LL2)

```
> Lower figures are better. We keep the LL3 model.

## Classical time estimation

Estimation of T05, T50 and T95:

```{r}

ED(mod_05_LL3, c(5,50,95), interval="delta")

```

Coefficient of variation of the different parameters in percent of the measure.
Lower is better:

```{r}

ED(mod_05_LL2,c(50),display=FALSE)[,2]*100/ED(mod_05_LL2,c(50),display=FALSE)[,1]
ED(mod_05_LL2,c(5),display=FALSE)[,2]*100/ED(mod_05_LL2,c(5),display=FALSE)[,1]
ED(mod_05_LL2,c(95),display=FALSE)[,2]*100/ED(mod_05_LL2,c(95),display=FALSE)[,1]

```

The best estimate is for T50.

## Comparison of the parameters

Direct model parameter comparison using the `compParm()` function. We use the
name we defined previously.

### Maximum germination rate

Comparison of the maximum rate of germination

```{r Max_comp_05}

compParm(mod_05_LL3, "Max")

```

No statistical difference. A lower rate of germination almost significant for RNAi 9.2.

### Slope

We compare the "Slope" that is the "vigor" of germination. 

```{r slope_comp_05}

slope_comp_05 <-
    compParm(mod_05_LL3, "Slope")

```
We have a small difference with Col compared to the others. We use multcompLetters to build the letter group after
adjusting for multi-testing (FDR)

```{r slope_comp_letter_05}

slope_comp_letters_05 <- 
    slope_comp_05[,"p-value"] %>%
    p.adjust(method="fdr") %>%
    setNames(str_replace(names(.), "/", "-")) %>%
    multcompLetters
slope_comp_letters_05$Letters

slope_comp_letters_05 <- 
    slope_comp_05[,"p-value"] %>%
    setNames(str_replace(names(.), "/", "-")) %>%
    multcompLetters
slope_comp_letters_05$Letters

```
If we correct with FDR, no difference.
Combine the total for a nice graph for variation of slope.

```{r slope_stats_05}

slope_stats_05 <-
    coef(summary(mod_05_LL2))[,1:2] %>%
    as.data.frame %>%
    rownames_to_column("parameters") %>%
    separate(parameters, c("parameter", "genotype"), sep=':') %>%
    filter(parameter == "Slope") %>%
    dplyr::select(genotype, slope=Estimate, se=`Std. Error`) %>%
    left_join(slope_comp_letters_05$Letters %>%
              as.data.frame %>%
              setNames("letters") %>%
              rownames_to_column("genotype")) %>%
    mutate(slope=-slope) %>%
    left_join(genotype_info)

```

And the graph itself

```{r}

shift <- max(slope_stats_05$se + slope_stats_05$slope) * .05

ggplot(data = slope_stats_05, aes(x = genotype_name, y = slope, fill=genotype_name)) +
    geom_col() +
    geom_errorbar(aes(ymin = slope - se, ymax = slope + se), width=.2) +
    geom_text(aes(y=slope + se + shift, label=letters)) +
    scale_fill_brewer(palette="Set2") +
    xlab("") + ylab("Speed of germination at T50 (proportion/hour)") +
    ggtitle("0.5 ABA")

```

Interestingly the mutation of *MHAL* seems to decrease the slope (not significant).

### T50

We then test T50, that it the time for the seeds to germinate at 50%.

```{r t50_comp_05}

T50_comp_05 <-
    compParm(mod_05_LL3, "T50")

```

Here are the real differences.
- Col compared to every body
- mrn1 mutant different from the rest
- no difference between the 2 RNAi and the 35S:MRN1

Other differences exist. We use `multcompLetters` to build the letter group after
adjusting for multi-testing (FDR)

```{r t50_comp_letter_05}

T50_comp_letters_05 <- 
    T50_comp_05[,"p-value"] %>%
    p.adjust(method="fdr") %>%
    setNames(str_replace(names(.), "/", "-")) %>%
    multcompLetters
T50_comp_letters_05

```

Combine the total for a nice graph for variation of T50.

```{r T50_stats_05}

T50_stats_05 <-
    coef(summary(mod_05_LL3))[,1:2] %>%
    as.data.frame %>%
    rownames_to_column("parameters") %>%
    separate(parameters, c("parameter", "genotype"), sep=':') %>%
    filter(parameter == "T50") %>%
    dplyr::select(genotype, T50=Estimate, se=`Std. Error`) %>%
    left_join(T50_comp_letters_05$Letters %>%
              as.data.frame %>%
              setNames("letters") %>%
              rownames_to_column("genotype")) %>%
    left_join(genotype_info)

```

And the graph itself

```{r}

shift <- max(T50_stats_05$se + T50_stats_05$T50) * .05

ggplot(data = T50_stats_05, aes(x = genotype_name, y = T50, fill=genotype_name)) +
    geom_col() +
    geom_errorbar(aes(ymin = T50 - se, ymax = T50 + se), width=.2) +
    geom_text(aes(y=T50 + se + shift, label=letters)) +
    scale_fill_brewer(palette="Set2") +
    xlab("") + ylab("Time to 50% of germination (hour)") +
    ggtitle("0.5 ABA")

```

# 0 ABA

## 3 parameters model fitting

```{r LL3_00, error=TRUE}

mod_00_LL3 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             curveid = genotype,
             data = formated_data,
             fct = LL.3(names=c("Slope", "Max", "T50")),
             type = "event", subset=c(concentration == 0)
    )

```
The classical logistic model failed to fit. Certainly because the germination
occur too fast. We can use the `Weibull` function that at the difference of the
logistic model is not symmetric compared to the inflexion point (T50).

Two types of model, first type 1

```{r W13_00, error=TRUE}

mod_00_W13 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             curveid = genotype,
             data = formated_data,
             fct = W1.3(names=c("Slope", "Max", "T50")),
             type = "event", subset=c(concentration == 0),
             upperl=c(NA,1,NA)
    )

```

Still failed, let try the type 2.

```{r W23_00, error=TRUE}

mod_00_W23 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             curveid = genotype,
             data = formated_data,
             fct = W2.3(names=c("Slope", "Max", "T50")),
             type = "event", subset=c(concentration == 0)
    )

```
No change no possibility to fit a model with 3 parameters (variable Maximum of
germination).

## 2 parameters model fitting

We will try with a 2 parameters model were the maximum is fixed and set at 1.
In a 2 parameter the max is set to 1.

```{r LL2_00, error=TRUE}

mod_00_LL2 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             curveid = genotype,
             data = formated_data,
             fct = LL.2(names=c("Slope", "T50")),
             type = "event", subset=c(concentration == 0))

```

The logistic model failed. We try the Weibull type 1

```{r W12_00, error=TRUE}

mod_00_W12 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             curveid = genotype,
             data = formated_data,
             fct = W1.2(names=c("Slope", "T50")),
             type = "event", subset=c(concentration == 0))

```

Failed again, so we test the Weibull type 2 model.

```{r W22_00}

mod_00_W22 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             curveid = genotype,
             data = formated_data,
             fct = W2.2(names=c("Slope", "T50")),
             type = "event", subset=c(concentration == 0))
plot(mod_00_W22, log = "", legendPos = c(80, .5),
     ylab="Proportion of germinated seed", xlab="Time (h)", main="0.5 ABA")
summary(mod_00_W22)

```

## Global effect

Again we check the difference with the null model.

```{r modNull_00_W22, error=TRUE}

modNull_00_W22 <- 
    drc::drm(DailyGerm ~ timeBef + timeAf,
             data = formated_data,
             fct = W2.2(names=c("Slope", "T50")),
             type = "event", subset=c(concentration == 0))

```

The null model failed. Let continue.

## Classical time estimation


Estimation of T05, T50 and T95:

```{r}

ED(mod_00_W22, c(5,50,95), interval="delta")

```

Coefficient of variation of the different parameters in percent of the mesure.
Lower is better:

```{r}

ED(mod_00_W22,c(50),display=FALSE)[,2]*100/ED(mod_00_W22,c(50),display=FALSE)[,1]
ED(mod_00_W22,c(5),display=FALSE)[,2]*100/ED(mod_00_W22,c(5),display=FALSE)[,1]
ED(mod_00_W22,c(95),display=FALSE)[,2]*100/ED(mod_00_W22,c(95),display=FALSE)[,1]

```

This time the better estimate is at 95% of germination. To have better estimates
we will need to have a better time resolution kinetic, especially in the short
times.

## Comparison of the parameters

Direct model parameter comparison using the `compParm()` function. We use the
name we defined previously.

### Slope

We compare the "Slope" that is the "vigor" of germination. 

```{r slope_comp_00}

slope_comp_00 <-
    compParm(mod_00_W22, "Slope")

```
This time it seems that the difference is with RNAi 3.8. We use multcompLetters to build the letter group after
adjusting for multi-testing (FDR)

```{r slope_comp_letter_00}

slope_comp_letters_00 <- 
    slope_comp_00[,"p-value"] %>%
    p.adjust(method="fdr") %>%
    setNames(str_replace(names(.), "/", "-")) %>%
    multcompLetters
slope_comp_letters_00$Letters

```
Combine the total for a nice graph for variation of slope.

```{r slope_stats_00}

slope_stats_00 <-
    coef(summary(mod_00_LL2))[,1:2] %>%
    as.data.frame %>%
    rownames_to_column("parameters") %>%
    separate(parameters, c("parameter", "genotype"), sep=':') %>%
    filter(parameter == "Slope") %>%
    dplyr::select(genotype, slope=Estimate, se=`Std. Error`) %>%
    left_join(slope_comp_letters_00$Letters %>%
              as.data.frame %>%
              setNames("letters") %>%
              rownames_to_column("genotype")) %>%
    left_join(genotype_info)

```

And the graph itself

```{r}

shift <- max(slope_stats_00$se + slope_stats_00$slope) * .05

ggplot(data = slope_stats_00, aes(x = genotype_name, y = slope, fill=genotype_name)) +
    geom_col() +
    geom_errorbar(aes(ymin = slope - se, ymax = slope + se), width=.2) +
    geom_text(aes(y=slope + se + shift, label=letters)) +
    scale_fill_brewer(palette="Set2") +
    xlab("") + ylab("Speed of germination at T50 (proportion/hour)") +
    ggtitle("0 ABA")

```

### T50

We then test T50, that it the time for the seeds to germination

Finally the T50, that it the time for the seeds to germinate at 50%.

```{r t50_comp_00}

T50_comp_00 <-
    compParm(mod_00_LL2, "T50")

```

Again difference with Col and the different RNAi. We use `multcompLetters` to build the letter group after
adjusting for multi-testing (FDR)

```{r t50_comp_letter_00}

T50_comp_letters_00 <- 
    T50_comp_00[,"p-value"] %>%
    p.adjust(method="fdr") %>%
    setNames(str_replace(names(.), "/", "-")) %>%
    multcompLetters
T50_comp_letters_00

```

Combine the total for a nice graph for variation of T50.

```{r T50_stats_00}

T50_stats_00 <-
    coef(summary(mod_00_W22))[,1:2] %>%
    as.data.frame %>%
    rownames_to_column("parameters") %>%
    separate(parameters, c("parameter", "genotype"), sep=':') %>%
    filter(parameter == "T50") %>%
    dplyr::select(genotype, T50=Estimate, se=`Std. Error`) %>%
    left_join(T50_comp_letters_00$Letters %>%
              as.data.frame %>%
              setNames("letters") %>%
              rownames_to_column("genotype")) %>%
    left_join(genotype_info)

```

And the graph itself

```{r}

shift <- max(T50_stats_00$se + T50_stats_00$T50) * .05

ggplot(data = T50_stats_00, aes(x = genotype_name, y = T50, fill=genotype_name)) +
    geom_col() +
    geom_errorbar(aes(ymin = T50 - se, ymax = T50 + se), width=.2) +
    geom_text(aes(y=T50 + se + shift, label=letters)) +
    scale_fill_brewer(palette="Set2") +
    xlab("") + ylab("Time to 50% of germination (hour)") +
    ggtitle("0.5 ABA")

```

# References
